---
title: "censusapi"
author: "Matt Colantonio"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
rm(list = ls()) 
gc()
cat("\f")
```

Loading the necessary packages

```{r}
library(tidyverse)
library(readr)
library(censusapi)
library(tidycensus)
```

This dataset contains parcel data for the City of Pittsburgh. It includes which census tract and census block each parcel is in. I want to download some demographic, housing, and economic data from the American Community Survey (ACS) to add to my parcel dataset.

```{r}
parcel_data <- read.csv("/Users/matthewcolantonio/Documents/Research/condemned_properties/residential_properties_pgh.csv")
```

Now that the dataset is loaded in, I want to create a list of census blocks included so I can extract data for these blocks using the Census API.

```{r}
# creating list of unique census blocks
unique_census_blocks <- unique(parcel_data$census_block)
# Convert the unique census blocks to a list
census_block_list <- as.list(unique_census_blocks)

# creating a list of unique census tracts
unique_census_tracts <- unique(parcel_data$census_tract)
# Convert the unique census blocks to a list
census_tract_list <- as.list(unique_census_tracts)

```

Now add your unique the Census API key.

```{r include=FALSE}
# Add key to .Renviron
Sys.setenv(CENSUS_KEY="YOUR KEY HERE")
# Reload .Renviron
readRenviron("~/.Renviron")
# Check to see that the expected key is output in your R console
Sys.getenv("CENSUS_KEY")

```

```{r}
# Variables and geography levels available in ACS 5-yr subject dataset as of 2021
head(listCensusMetadata(name = "acs/acs5/subject", vintage = 2021, type = "variables"))
head(listCensusMetadata(name = "acs/acs5/subject", vintage = 2021, type = "geography"))

head(listCensusMetadata(name = "acs/acs5/profile", vintage = 2021, type = "variables"))
head(listCensusMetadata(name = "acs/acs5/profile", vintage = 2021, type = "geography"))
```

```{r message=FALSE, warning=FALSE}

# Set your Census API key
census_api_key("e8c3122e8584f0de407bfdd90c617f2213408ffe")

# Define the variables you want to retrieve from the Economic Characteristics dataset
variables <- c("DP03_0002E", "DP03_0003E", "DP03_0004E", "DP03_0005E")

# Set your desired year
year <- 2019

# Set the geography, which is 'tract' in this case
geography <- "tract"

# Specify the state and county (Allegheny, Pennsylvania)
state <- "PA"
county <- "Allegheny"

# Filter the unique census tracts for Allegheny, Pennsylvania
filtered_census_tracts <- unique_census_tracts  # Replace with your actual list of census tracts

# Create an empty data frame to store the results
economic_data <- data.frame()

# Loop through the filtered census tracts and fetch data
for (tract in filtered_census_tracts) {
  data <- get_acs(geography = geography, variables = variables, 
                  year = year, state = state, county = county, tract = tract)
  
  economic_data <- bind_rows(economic_data, data)
}

# Now, 'economic_data' contains the Economic Characteristics data for the specified census tracts in Allegheny, Pennsylvania

```

```{r}
head(economic_data)
```
