---
title: "County Level Data for Appalachia"
author: "Matt Colantonio"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
toc_depth: 3
---

```{r message=FALSE, warning=FALSE}
# this code clears the environment and installs/loads popular packages

rm(list = ls()) 
  gc()            
  cat("\f")  
  
packages <- c("readr", #open csv
              "psych", # quick summary stats for data exploration,
              "stargazer", #summary stats for sharing,
              "tidyverse", # data manipulation like selecting variables,
              "corrplot", # correlation plots
              "ggplot2", # graphing
              "ggcorrplot", # correlation plot
              "gridExtra", #overlay plots
              "data.table", # reshape for graphing 
              "car", #vif
              "prettydoc", # html output
              "visdat", # visualize missing variables
              "glmnet", # lasso/ridge
              "caret", # confusion matrix
              "MASS", #step AIC
              "plm", # fixed effects demeaned regression
              "lmtest", # test regression coefficients
              "fpp3", # Foprecasting: Principles & Practice supplement
              "tsibble", 
              "tsibbledata",
              "lubridate",
              "forecast", # time sereis forecasting
              "bea.R" # interact with BEA APISs
)

for (i in 1:length(packages)) {
  if (!packages[i] %in% rownames(installed.packages())) {
    install.packages(packages[i]
                     , repos = "http://cran.rstudio.com/"
                     , dependencies = TRUE
    )
  }
  library(packages[i], character.only = TRUE)
}

rm(packages)
```

```{r}
setwd("/Users/matthewcolantonio/Documents/Research/fracking/")
```

```{r}
beaKey <-  "D84279EC-396F-4A6F-BE18-D9279CD3B1B0"
```

```{r}
appalachia <- list(
  "Alabama" = c("Bibb, AL", "Blount, AL", "Calhoun, AL", "Chambers, AL", "Cherokee, AL", "Chilton, AL", "Clay, AL", "Cleburne, AL", "Colbert, AL", "Coosa, AL", "Cullman, AL", "DeKalb, AL", "Elmore, AL", "Etowah, AL", "Fayette, AL", "Franklin, AL", "Hale, AL", "Jackson, AL", "Jefferson, AL", "Lamar, AL", "Lauderdale, AL", "Lawrence, AL", "Limestone, AL", "Macon, AL", "Madison, AL", "Marion, AL", "Marshall, AL", "Morgan, AL", "Pickens, AL", "Randolph, AL", "St. Clair, AL", "Shelby, AL", "Talladega, AL", "Tallapoosa, AL", "Tuscaloosa, AL", "Walker, AL", "Winston, AL"),
  "Georgia" = c("Banks, GA", "Barrow, GA", "Bartow, GA", "Carroll, GA", "Catoosa, GA", "Chattooga, GA", "Cherokee, GA", "Dade, GA", "Dawson, GA", "Douglas, GA", "Elbert, GA", "Fannin, GA", "Floyd, GA", "Forsyth, GA", "Franklin, GA", "Gilmer, GA", "Gordon, GA", "Gwinnett, GA", "Habersham, GA", "Hall, GA", "Haralson, GA", "Hart, GA", "Heard, GA", "Jackson, GA", "Lumpkin, GA", "Madison, GA", "Murray, GA", "Paulding, GA", "Pickens, GA", "Polk, GA", "Rabun, GA", "Stephens, GA", "Towns, GA", "Union, GA", "Walker, GA", "White, GA", "Whitfield, GA"),
  "Kentucky" = c("Adair, KY", "Bath, KY", "Bell, KY", "Boyd, KY", "Breathitt, KY", "Carter, KY", "Casey, KY", "Clark, KY", "Clay, KY", "Clinton, KY", "Cumberland, KY", "Edmonson, KY", "Elliott, KY", "Estill, KY", "Fleming, KY", "Floyd, KY", "Garrard, KY", "Green, KY", "Greenup, KY", "Harlan, KY", "Hart, KY", "Jackson, KY", "Johnson, KY", "Knott, KY", "Knox, KY", "Laurel, KY", "Lawrence, KY", "Lee, KY", "Leslie, KY", "Letcher, KY", "Lewis, KY", "Lincoln, KY", "McCreary, KY", "Madison, KY", "Magoffin, KY", "Martin, KY", "Menifee, KY", "Metcalfe, KY", "Monroe, KY", "Montgomery, KY", "Morgan, KY", "Nicholas, KY", "Owsley, KY", "Perry, KY", "Pike, KY", "Powell, KY", "Pulaski, KY", "Robertson, KY", "Rockcastle, KY", "Rowan, KY", "Russell, KY", "Wayne, KY", "Whitley, KY", "Wolfe, KY"),
  "Maryland" = c("Allegany, MD", "Garrett, MD", "Washington, MD"),
  "Mississippi" = c("Alcorn, MS", "Benton, MS", "Calhoun, MS", "Chickasaw, MS", "Choctaw, MS", "Clay, MS", "Itawamba, MS", "Kemper, MS", "Lee, MS", "Lowndes, MS", "Marshall, MS", "Monroe, MS", "Montgomery, MS", "Noxubee, MS", "Oktibbeha, MS", "Panola, MS", "Pontotoc, MS", "Prentiss, MS", "Tippah, MS", "Tishomingo, MS", "Union, MS", "Webster, MS", "Winston, MS", "Yalobusha, MS"),
  "New York" = c("Allegany, NY", "Broome, NY", "Cattaraugus, NY", "Chautauqua, NY", "Chemung, NY", "Chenango, NY", "Cortland, NY", "Delaware, NY", "Otsego, NY", "Schoharie, NY", "Schuyler, NY", "Steuben, NY", "Tioga, NY", "Tompkins, NY"),
  "North Carolina" = c("Alexander, NC", "Alleghany, NC", "Ashe, NC", "Avery, NC", "Buncombe, NC", "Burke, NC", "Caldwell, NC", "Catawba, NC", "Cherokee, NC", "Clay, NC", "Cleveland, NC", "Davie, NC", "Forsyth, NC", "Graham, NC", "Haywood, NC", "Henderson, NC", "Jackson, NC", "McDowell, NC", "Macon, NC", "Madison, NC", "Mitchell, NC", "Polk, NC", "Rutherford, NC", "Stokes, NC", "Surry, NC", "Swain, NC", "Transylvania, NC", "Watauga, NC", "Wilkes, NC", "Yadkin, NC", "Yancey, NC"),
  "Ohio" = c("Adams, OH", "Ashtabula, OH", "Athens, OH", "Belmont, OH", "Brown, OH", "Carroll, OH", "Clermont, OH", "Columbiana, OH", "Coshocton, OH", "Gallia, OH", "Guernsey, OH", "Harrison, OH", "Highland, OH", "Hocking, OH", "Holmes, OH", "Jackson, OH", "Jefferson, OH", "Lawrence, OH", "Mahoning, OH", "Meigs, OH", "Monroe, OH", "Morgan, OH", "Muskingum, OH", "Noble, OH", "Perry, OH", "Pike, OH", "Ross, OH", "Scioto, OH", "Trumbull, OH", "Tuscarawas, OH", "Vinton, OH", "Washington, OH"),
  "Pennsylvania" = c("Allegheny, PA", "Armstrong, PA", "Beaver, PA", "Bedford, PA", "Blair, PA", "Bradford, PA", "Butler, PA", "Cambria, PA", "Cameron, PA", "Carbon, PA", "Centre, PA", "Clarion, PA", "Clearfield, PA", "Clinton, PA", "Columbia, PA", "Crawford, PA", "Elk, PA", "Erie, PA", "Fayette, PA", "Forest, PA", "Fulton, PA", "Greene, PA", "Huntingdon, PA", "Indiana, PA", "Jefferson, PA", "Juniata, PA", "Lackawanna, PA", "Lawrence, PA", "Luzerne, PA", "Lycoming, PA", "McKean, PA", "Mercer, PA", "Mifflin, PA", "Monroe, PA", "Montour, PA", "Northumberland, PA", "Perry, PA", "Pike, PA", "Potter, PA", "Schuylkill, PA", "Snyder, PA", "Somerset, PA", "Sullivan, PA", "Susquehanna, PA", "Tioga, PA", "Union, PA", "Venango, PA", "Warren, PA", "Washington, PA", "Wayne, PA", "Westmoreland, PA", "Wyoming, PA"),
  "South Carolina" = c("Anderson, SC", "Cherokee, SC", "Greenville, SC", "Oconee, SC", "Pickens, SC", "Spartanburg, SC", "Union, SC"),
  "Tennessee" = c("Anderson, TN", "Bledsoe, TN", "Blount, TN", "Bradley, TN", "Campbell, TN", "Cannon, TN", "Carter, TN", "Claiborne, TN", "Clay, TN", "Cocke, TN", "Coffee, TN", "Cumberland, TN", "DeKalb, TN", "Fentress, TN", "Franklin, TN", "Grainger, TN", "Greene, TN", "Grundy, TN", "Hamblen, TN", "Hamilton, TN", "Hancock, TN", "Hawkins, TN", "Jackson, TN", "Jefferson, TN", "Johnson, TN", "Knox, TN", "Lawrence, TN", "Lewis, TN", "Loudon, TN", "McMinn, TN", "Macon, TN", "Marion, TN", "Meigs, TN", "Monroe, TN", "Morgan, TN", "Overton, TN", "Pickett, TN", "Polk, TN", "Putnam, TN", "Rhea, TN", "Roane, TN", "Scott, TN", "Sequatchie, TN", "Sevier, TN", "Smith, TN", "Sullivan, TN", "Unicoi, TN", "Union, TN", "Van Buren, TN", "Warren, TN", "Washington, TN", "White, TN"),
  "Virginia" = c("Alleghany, VA", "Bath, VA", "Bland, VA", "Botetourt, VA", "Buchanan, VA", "Carroll, VA", "Craig, VA", "Dickenson, VA", "Floyd, VA", "Giles, VA", "Grayson, VA", "Henry, VA", "Highland, VA", "Lee, VA", "Montgomery, VA", "Patrick, VA", "Pulaski, VA", "Rockbridge, VA", "Russell, VA", "Scott, VA", "Smyth, VA", "Tazewell, VA", "Washington, VA", "Wise, VA", "Wythe, VA"),
  "Virginia Cities" = c("Bristol, VA", "Buena Vista, VA", "Covington, VA", "Galax, VA", "Lexington, VA", "Martinsville, VA", "Norton, VA", "Radford, VA"),
  "West Virginia" = c("Barbour, WV", "Berkeley, WV", "Boone, WV", "Braxton, WV", "Brooke, WV", "Cabell, WV", "Calhoun, WV", "Clay, WV", "Doddridge, WV", "Fayette, WV", "Gilmer, WV", "Grant, WV", "Greenbrier, WV", "Hampshire, WV", "Hancock, WV", "Hardy, WV", "Harrison, WV", "Jackson, WV", "Jefferson, WV", "Kanawha, WV", "Lewis, WV", "Lincoln, WV", "Logan, WV", "Marion, WV", "Marshall, WV", "Mason, WV", "McDowell, WV", "Mercer, WV", "Mineral, WV", "Mingo, WV", "Monongalia, WV", "Monroe, WV", "Morgan, WV", "Nicholas, WV", "Ohio, WV", "Pendleton, WV", "Pleasants, WV", "Pocahontas, WV", "Preston, WV", "Putnam, WV", "Raleigh, WV", "Randolph, WV", "Ritchie, WV", "Roane, WV", "Summers, WV", "Taylor, WV", "Tucker, WV", "Tyler, WV", "Upshur, WV", "Wayne, WV", "Webster, WV", "Wetzel, WV", "Wirt, WV", "Wood, WV", "Wyoming, WV")
)

```

## CAGGDP9 Real GDP in Chained Dollars

### All Industry total (1)

```{r}
gdp1 <- list(
  "UserID" = beaKey, 
  "Method" = "GetData", 
  "datasetname" = "Regional", 
  "TableName" = "CAGDP9", 
  "LineCode" = 1,  
  "GeoFips" = "COUNTY",
  "Year" = "LAST5"
)

gdp1 <- beaGet(gdp1, asWide = FALSE)
```

```{r}
is_in_appalachia <- lapply(gdp1$GeoName, function(county) {
  any(county %in% unlist(appalachia))
})

# Subset your dataset to include only the desired counties
gdp1.1 <- gdp1[unlist(is_in_appalachia), ]
```

### Mining, quarrying, and oil and gas extraction (6)

```{r}
gdp6 <- list(
  "UserID" = beaKey, 
  "Method" = "GetData", 
  "datasetname" = "Regional", 
  "TableName" = "CAGDP9", 
  "LineCode" = 6,  
  "GeoFips" = "COUNTY",
  "Year" = "LAST5"
)

gdp6 <- beaGet(gdp6, asWide = FALSE)

is_in_appalachia <- lapply(gdp6$GeoName, function(county) {
  any(county %in% unlist(appalachia))
})

# Subset your dataset to include only the desired counties
gdp6.1 <- gdp6[unlist(is_in_appalachia), ]
```

### Natural resources and mining (87)

```{r}
gdp87 <- list(
  "UserID" = beaKey, 
  "Method" = "GetData", 
  "datasetname" = "Regional", 
  "TableName" = "CAGDP9", 
  "LineCode" = 87,  
  "GeoFips" = "COUNTY",
  "Year" = "LAST5"
)

gdp87 <- beaGet(gdp87, asWide = FALSE)

is_in_appalachia <- lapply(gdp87$GeoName, function(county) {
  any(county %in% unlist(appalachia))
})

# Subset your dataset to include only the desired counties
gdp87.1 <- gdp87[unlist(is_in_appalachia), ]
```

### GDP Merged

```{r}
# rename columns
"AllIndustryGDP" -> colnames(gdp1)[7] 
"AllIndustryGDP" -> colnames(gdp1.1)[7] 

"extractionGDP" -> colnames(gdp6)[7] 
"extractionGDP" -> colnames(gdp6.1)[7]

"nat_resourceGDP" -> colnames(gdp87)[7]
"nat_resourceGDP" -> colnames(gdp87.1)[7]
```

```{r}
# Appalachia
# Merge gdp1.1, gdp6.1, and gdp87.1 by GeoName and TimePeriod
merged_gdp.1 <- merge(gdp1.1, gdp6.1, by = c("GeoName", "TimePeriod"))
merged_gdp.1 <- merge(merged_gdp.1, gdp87.1, by = c("GeoName", "TimePeriod"))

merged_gdp.1 <- merged_gdp.1[, c("GeoName", "TimePeriod", "AllIndustryGDP", "extractionGDP", "nat_resourceGDP")
]
```

```{r}
# US total
# Merge gdp1, gdp6, and gdp87 by GeoName and TimePeriod
merged_gdp <- merge(gdp1, gdp6, by = c("GeoName", "TimePeriod"))
merged_gdp <- merge(merged_gdp, gdp87, by = c("GeoName", "TimePeriod"))

merged_gdp <- merged_gdp[, c("GeoName", "TimePeriod", "AllIndustryGDP", "extractionGDP", "nat_resourceGDP")
]
```

```{r}
# we can remove all former dataframes to keep the environment clean
# merged_gdp / merged_gdp.1 datasets currently contain all counties for county-level analysis or GIS mapping

rm(gdp1, gdp1.1, gdp6, gdp6.1, gdp87, gdp87.1)
```

Now there are two datasets containing GDP data: one for all counties in the US and one for Appalachian counties only.

We can further filter our data:

I want to remove Appalachian counties from the total US data frame so GDP calculations can accurately measure the impact of Appalachia to extraction and natural resource GDP.

```{r}
merged_gdp_filtered <- merged_gdp %>%
  filter(!GeoName %in% unlist(appalachia))
```

### Calculations

```{r}
# percent of GDP in exctraction of oil and gas/mining Appalachia
sum(merged_gdp.1$extractionGDP[merged_gdp.1$TimePeriod == 2021]) /
sum(merged_gdp.1$AllIndustryGDP[merged_gdp.1$TimePeriod == 2021])
```

```{r}
# percent of GDP in exctraction of oil and gas/mining US
sum(merged_gdp$extractionGDP[merged_gdp$TimePeriod == 2021]) /
sum(merged_gdp$AllIndustryGDP[merged_gdp$TimePeriod == 2021])
```

```{r}
# percent of GDP in exctraction of oil and gas/mining Appalachia
sum(merged_gdp.1$nat_resourceGDP[merged_gdp.1$TimePeriod == 2021]) /
sum(merged_gdp.1$AllIndustryGDP[merged_gdp.1$TimePeriod == 2021])

# percent of GDP in exctraction of oil and gas/mining US
sum(merged_gdp$nat_resourceGDP[merged_gdp$TimePeriod == 2021]) /
sum(merged_gdp$AllIndustryGDP[merged_gdp$TimePeriod == 2021])
```

```{r}
# extraction + natural resources
# Appalachia
(sum(merged_gdp.1$nat_resourceGDP[merged_gdp.1$TimePeriod == 2021]) + sum(merged_gdp.1$extractionGDP[merged_gdp.1$TimePeriod == 2021])) /
sum(merged_gdp.1$AllIndustryGDP[merged_gdp.1$TimePeriod == 2021])


(sum(merged_gdp$nat_resourceGDP[merged_gdp$TimePeriod == 2021]) + sum(merged_gdp$extractionGDP[merged_gdp$TimePeriod == 2021])) /
sum(merged_gdp$AllIndustryGDP[merged_gdp$TimePeriod == 2021])
```

```{r}
# percentage of natural resources and extraction GDP to total US GDP in those industries
(sum(merged_gdp.1$nat_resourceGDP[merged_gdp.1$TimePeriod == 2021]) + sum(merged_gdp.1$extractionGDP[merged_gdp.1$TimePeriod == 2021])) /
(sum(merged_gdp$nat_resourceGDP[merged_gdp$TimePeriod == 2021]) + sum(merged_gdp$extractionGDP[merged_gdp$TimePeriod == 2021]))


(sum(merged_gdp_filtered$nat_resourceGDP[merged_gdp_filtered$TimePeriod == 2021]) + sum(merged_gdp_filtered$extractionGDP[merged_gdp_filtered$TimePeriod == 2021])) /
(sum(merged_gdp$nat_resourceGDP[merged_gdp$TimePeriod == 2021]) + sum(merged_gdp$extractionGDP[merged_gdp$TimePeriod == 2021]))



```

## CAINC30 Economic Profile

### Population (100)

```{r}
ep100 <- list(
  "UserID" = beaKey, 
  "Method" = "GetData", 
  "datasetname" = "Regional", 
  "TableName" = "CAINC30", 
  "LineCode" = 100,  
  "GeoFips" = "COUNTY",
  "Year" = "LAST5"
)

ep100 <- beaGet(ep100, asWide = FALSE)

is_in_appalachia <- lapply(ep100$GeoName, function(county) {
  any(county %in% unlist(appalachia))
})

# Subset your dataset to include only the desired counties
ep100.1 <- ep100[unlist(is_in_appalachia), ]
```

### Total Employment (240)

```{r}
ep240 <- list(
  "UserID" = beaKey, 
  "Method" = "GetData", 
  "datasetname" = "Regional", 
  "TableName" = "CAINC30", 
  "LineCode" = 240,  
  "GeoFips" = "COUNTY",
  "Year" = "LAST5"
)

ep240 <- beaGet(ep240, asWide = FALSE)

is_in_appalachia <- lapply(ep240$GeoName, function(county) {
  any(county %in% unlist(appalachia))
})

# Subset your dataset to include only the desired counties
ep240.1 <- ep240[unlist(is_in_appalachia), ]
```

### Personal Income (10)

```{r}
ep10 <- list(
  "UserID" = beaKey, 
  "Method" = "GetData", 
  "datasetname" = "Regional", 
  "TableName" = "CAINC30", 
  "LineCode" = 10,  
  "GeoFips" = "COUNTY",
  "Year" = "LAST5"
)

ep10 <- beaGet(ep10, asWide = FALSE)

is_in_appalachia <- lapply(ep10$GeoName, function(county) {
  any(county %in% unlist(appalachia))
})

# Subset your dataset to include only the desired counties
ep10.1 <- ep10[unlist(is_in_appalachia), ]
```

### Economic Profile Merged

```{r}
# rename columns
"population" -> colnames(ep100)[7] 
"population" -> colnames(ep100.1)[7] 

"total_employment" -> colnames(ep240)[7] 
"total_employment" -> colnames(ep240.1)[7]

"personal_income" -> colnames(ep10)[7]
"personal_income" -> colnames(ep10.1)[7]
```

```{r}
# Appalachia
# Merge gdp1.1, gdp6.1, and gdp87.1 by GeoName and TimePeriod
merged_ep.1 <- merge(ep100.1, ep240.1, by = c("GeoName", "TimePeriod"))
merged_ep.1 <- merge(merged_ep.1, ep10.1, by = c("GeoName", "TimePeriod"))

merged_ep.1 <- merged_ep.1[, c("GeoName", "TimePeriod", "population", "total_employment", "personal_income")
]
```

```{r}
# US
merged_ep <- merge(ep100, ep240, by = c("GeoName", "TimePeriod"))
merged_ep <- merge(merged_ep, ep10, by = c("GeoName", "TimePeriod"))

merged_ep <- merged_ep[, c("GeoName", "TimePeriod", "population", "total_employment", "personal_income")
]

# US less Appalachian counties
merged_ep_filtered <- merged_ep %>%
  filter(!GeoName %in% unlist(appalachia))
```

```{r}
rm(ep10, ep10.1, ep100, ep100.1, ep240, ep240.1)
```

## CAINC6N Compensation by employees by NAICS Industry

### Oil and gas extraction (201)

```{r}
compensation201 <- list(
  "UserID" = beaKey, 
  "Method" = "GetData", 
  "datasetname" = "Regional", 
  "TableName" = "CAINC6N", 
  "LineCode" = 201,  
  "GeoFips" = "COUNTY",
  "Year" = "LAST5"
)

compensation201 <- beaGet(compensation201, asWide = FALSE)

is_in_appalachia <- lapply(compensation201$GeoName, function(county) {
  any(county %in% unlist(appalachia))
})

# Subset your dataset to include only the desired counties
compensation201.1 <- compensation201[unlist(is_in_appalachia), ]
```

### Petroleum and coal products manufacturing (539)

```{r}
compensation539 <- list(
  "UserID" = beaKey, 
  "Method" = "GetData", 
  "datasetname" = "Regional", 
  "TableName" = "CAINC6N", 
  "LineCode" = 539,  
  "GeoFips" = "COUNTY",
  "Year" = "LAST5"
)

compensation539 <- beaGet(compensation539, asWide = FALSE)

is_in_appalachia <- lapply(compensation539$GeoName, function(county) {
  any(county %in% unlist(appalachia))
})

# Subset your dataset to include only the desired counties
compensation539.1 <- compensation539[unlist(is_in_appalachia), ]
```

### Pipeline transportation (806)

```{r}
compensation806 <- list(
  "UserID" = beaKey, 
  "Method" = "GetData", 
  "datasetname" = "Regional", 
  "TableName" = "CAINC6N", 
  "LineCode" = 806,  
  "GeoFips" = "COUNTY",
  "Year" = "LAST5"
)

compensation806 <- beaGet(compensation806, asWide = FALSE)

is_in_appalachia <- lapply(compensation806$GeoName, function(county) {
  any(county %in% unlist(appalachia))
})

# Subset your dataset to include only the desired counties
compensation806.1 <- compensation806[unlist(is_in_appalachia), ]
```

### Compensation Merged

```{r}
# rename columns
"oil_gas" -> colnames(compensation201)[7] 
"oil_gas" -> colnames(compensation201.1)[7] 

"petro_manufacturing" -> colnames(compensation539)[7] 
"petro_manufacturing" -> colnames(compensation539.1)[7]

"pipeline_transport" -> colnames(compensation806)[7]
"pipeline_transport" -> colnames(compensation806.1)[7]
```

```{r}
# Appalachia
merged_comp.1 <- merge(compensation201.1, compensation539.1, by = c("GeoName", "TimePeriod"))
merged_comp.1 <- merge(merged_comp.1, compensation806.1, by = c("GeoName", "TimePeriod"))

merged_comp.1 <- merged_comp.1[, c("GeoName", "TimePeriod", "oil_gas", "petro_manufacturing", "pipeline_transport")
]

```

```{r}
# US
merged_comp <- merge(compensation201, compensation539, by = c("GeoName", "TimePeriod"))
merged_comp <- merge(merged_comp, compensation806, by = c("GeoName", "TimePeriod"))

merged_comp <- merged_comp[, c("GeoName", "TimePeriod", "oil_gas", "petro_manufacturing", "pipeline_transport")
]

# US less Appalachian counties
merged_comp_filtered <- merged_comp %>%
  filter(!GeoName %in% unlist(appalachia))
```

```{r}
rm(compensation201, compensation201.1, compensation539, compensation539.1, compensation806, compensation806.1)
```

## Merging All

The ultimate goal is to have one data frame containing all the measures that were pulled from the BEA site using bea.R.

```{r}
# for all US counties
US <-  merge(merged_comp, merged_gdp, by = c("GeoName", "TimePeriod"))
US <- merge(US, merged_ep, by = c("GeoName", "TimePeriod"))
```

```{r}
# for Appalachian counties
app <-  merge(merged_comp.1, merged_gdp.1, by = c("GeoName", "TimePeriod"))
app <- merge(app, merged_ep.1, by = c("GeoName", "TimePeriod"))
```

```{r}
# for US counties NOT including Appalachia
filtered <-  merge(merged_comp_filtered, merged_gdp_filtered, by = c("GeoName", "TimePeriod"))
filtered <- merge(filtered, merged_ep_filtered, by = c("GeoName", "TimePeriod"))
```

Now we have three data frames:

1.  'app' contains data for Appalachian counties
2.  'US' contains data for all US counties (with available data)
3.  'filtered' contains data for US counties NOT including Appalachian counties

all for the last 5 years of available data, 2017-2021.

```{r}
# we can remove the separate data frames since we have merged them
rm(merged_comp, merged_comp.1, merged_comp_filtered, merged_ep, merged_ep.1, merged_ep_filtered, merged_gdp, merged_gdp.1, merged_gdp_filtered)
```

## Download Datasets

```{r}
# since the wd is set already, the files will appear there. if a wd is not set, write the file path in the code

write_csv(app, "app_data.csv")
write_csv(US, "US_data.csv")
write_csv(filtered, "filtered_data.csv")
```
